---
title: Merge data
output: github_document
---

The purpose of this document is to merge the data from all the sources into some useful tables.

```{r}
library("here")
library("vroom")
library("fs")
library("purrr")
library("dplyr")
library("conflicted")
library("readr")

conflict_prefer("filter", "dplyr")
```

Let's define the directories and create the target directory.

```{r}
dirs <- 
  list(
    source_state = path("data", "scrape-site"),
    source_nyt = path("data", "nyt"),
    target = path("data", "merged")  
  ) %>%
  map(here)

dir_create(dirs$target)
```

Let's read in the NYT data:

```{r}
nyt_data <- vroom(path(dirs$source_nyt, "nyt-iowa.csv"))
```

And the state data:

```{r}
state_data <- vroom(dir_ls(dirs$source_state))
```

Get the dates in the state dataset and exclude those from the NYT dataset.

```{r}
dates_state <- unique(state_data$date) %>% print() 
```

```{r}
nyt_data_abridged <- 
  nyt_data %>%
  filter(!date %in% dates_state)
```

Now we can bind the data frames together. I take into consideration:

- discarding rows unknown counties; these eventually get counted. 
- `cases`, `deaths`, `tests`, and `recovered` must always be increasing. Given a current value, all earlier values can be no greater than this value.
- `active_cases` is the number of `cases` less `deaths` and `recovered`

```{r}
merged <- 
  bind_rows(nyt_data_abridged, state_data) %>%
  arrange(desc(date), desc(cases)) %>%
  mutate(
    county = ifelse(is.na(fips), NA_character_, county)
  ) %>%
  group_by(county) %>%
  mutate(
    cases = cummin(cases),
    deaths = cummin(deaths),
    tests = cummin(tests),
    recovered = cummin(recovered),
    active_cases = cases - deaths - recovered
  ) %>%
  ungroup() %>%
  print()
```

Let's write this out:

```{r}
write_csv(
  merged,
  path(dirs$target, "merged.csv"),
  na = ""
)
```
